---
title: "Panther CLOGIT 50 pts"
author: "Wenjing Xu"
date: "4/17/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(dplyr)
library(ggplot2)
library(GGally)
library(arm)
#library(MCMCglmm)
#library(plotMCMC)
library(reshape)
#library(parallel)
library(aod)
library(fastDummies)
library(MuMIn)
library(hab)
library(cowplot)
```
This round using 50 random points instead of 10. 

## dataframe prep
```{r data prep}
# reclassify vegetation class
data <- read.csv("Final_Variable_all.csv")
veg.code <- read.csv("VegCombineCode.csv")
# reshape data
data <- data %>% dplyr::left_join(veg.code, by = "CLC_STATE")
data.full <- data %>% filter (!is.na(Fire_Cat))
# reduce to 10 random points for each den 
set.seed(17)
# decide random sample size (smaller than 100)
data.0 <- data.full %>% group_by (DenID) %>% filter( Used == 0 ) %>% sample_n(50)
data.1 <- data.full %>% filter( Used == 1 ) 
data <- union (data.1, data.0) %>% arrange (DenID)

# specify factors. Levels order will affect the baseline level in the modeling process
data$DenID <- as.factor(data$DenID)
data$Fire_Cat <- factor(data$Fire_Cat, levels = c("U", "A", "B", "C", "D"))
```

## Examining variables
```{r examining categorical variables, results="hide", include=FALSE}
# investigate vegetation class 
#data %>% filter (Used == "1") %>% group_by(Specific.Class) %>% summarise(used.count = n())
# what to do with Bare soil/clear cut and Exotic plants? -- need to examine neighboring class in maps and maybe reclassify based on that. For now, leave them out. 
#data %>% filter (Used == "1") %>% filter ((Specific.Class == "Bare Soil/Clear Cut") | (Specific.Class == "Exotic Plants")) %>% dplyr::select(DenID)
#data <- data %>% filter ((DenID != "61") & (DenID != "95A"))

# investiage combined class
#data %>% filter (Used == "1") %>% group_by(Combined) %>% summarise(used.count = n())
```


```{r reclass veg class}
data %>% group_by(Combined) %>% summarise(used.count = n())
# get rid of the factors that are ecologically unlikely to be meaningful
# or that likely do not have much explanatory power (drop CatID because not enough replicates for each cat)
# or are obviously correlated (e.g. ave_wet and ave_dry)
# also deleted age because does not really show importance but interfere with the modeling process from previous runs
data <- data %>% dplyr::select(Used, DenID, Fire_Cat, Ave_Wet, Prec_Tree, dist2build, Combined) %>% dplyr::rename(Veg_Class = Combined)
data$Veg_Class <- factor(data$Veg_Class, levels = c("Upland Forest", "Other", "Marsh-Shrub-Swamp", "Prairie-Grassland", "Wetland Forest"))
```

## data exploration

```{r outlier}
op <- par(mfrow = c(3, 1))
dotchart(data$Ave_Wet, main = "Average water height in wet season", group = data$Fire_Cat)
dotchart(data$dist2build, main = "Distance to Building", group = data$Fire_Cat)
dotchart(data$Prec_Tree, main = "Percept Tree", group = data$Fire_Cat)
par(op)
```
Age and dist2build has some outliers at the far end. So do transformation on these two. 

```{r colinearity}
data$L.dist2build <- log10(data$dist2build+1) #since some of the dist is 0
data.L <- data %>% dplyr::select(Used, DenID, Fire_Cat, Ave_Wet, Prec_Tree, L.dist2build, Veg_Class) 

cor(data.L[,c(4:6)]) #does not seem to have strong linear relations
ggpairs(data.L[, c( "Ave_Wet", "Prec_Tree", "L.dist2build")])
```

This plot further confirmed that there are no obvious linear relations. 

```{r categorical variables, include = FALSE}
#The table shows we have a unbalanced dataset in terms of Fire_Cat. In principle, lme4 can deal with unbalanced data sets but the low number of data points in some cells of the design means that it is hard to estimate some of the effects. 
data.L %>% group_by(Fire_Cat) %>% summarise(used.count = n())
par(mfrow = c(1, 2))
ggplot(data.L, aes(x = Fire_Cat, y = DenID)) +
  stat_sum(aes(size = ..n.., group = Fire_Cat, color = factor(Used))) +
  scale_size_area(max_size=10)
```
We have a lot of points in U. Unbalanced dataset. Later for modeling we will create one dataset that combines A and B, and C and D. 

```{r}
data.2 <- melt(data.L[, c("Used", "Ave_Wet", "Prec_Tree", "L.dist2build")], id.vars="Used")
ggplot(data.2, aes(factor(Used), y = value, fill=factor(Used))) +
    geom_boxplot() +
    facet_wrap(~variable, scales="free_y")

```
Super clear that they pick high tree precentage area. slightly far away from building, almost no difference in age and ave wet. We expect precentage tree will play an important role.

```{r scaling, include = FALSE}
# # scaling numeric variables by deviding two standard deviations for binary predictors
# # http://www.stat.columbia.edu/~gelman/research/published/standardizing7.pdf
# data.scaled <- data
# data.scaled$dist2build <- log(data.scaled$dist2build + 1)
# data.scaled[,5:8] <- sapply(data.scaled[,5:8], rescale, binary.inputs = "full")
# #data.scaled$Ave_Wet <- sqrt(data.scaled$Ave_Wet)
# #data.scaled$Prec_Tree <- sqrt(data.scaled$Prec_Tree)  # issue: take sqrt creates lots of NA 
# ggpairs(data.scaled[, c("Age", "Ave_Wet", "Prec_Tree", "dist2build")])
```


## fitting conditional logistic regression model and model selection
The model selection process can be described as evaluating candidate models (with Fire_cat always be one variable). We considered relevant interactions into the most parsimonious models to see if they would increase fit, but models with interactions always ranked poorer (higher AICc) than models excluding them. 
```{r fitting round 1}
# full model
m.0 <- clogit(Used ~ Fire_Cat + Prec_Tree + Veg_Class + Ave_Wet + L.dist2build + strata(DenID), data = data.L, method='efron')
summary(m.0)
# # if combine fire cat 
# data.LC <- data.L %>% mutate(Fire_Cat_C = ifelse (Fire_Cat == "A" | Fire_Cat == "B", "AB", 
#                              ifelse (Fire_Cat == "U", "U", "CD")))
# data.LC$Fire_Cat_C <- factor(data.LC$Fire_Cat_C, levels = c("AB", "CD", "U"))
# m.0c <- clogit(Used ~ Fire_Cat_C + Prec_Tree + Veg_Class + Ave_Wet + L.dist2build + strata(DenID), data = data.LC, method='efron')
# summary(m.0c)
# # combining class attenuate impacts of fire class C... decide to go back to not combined classes

m.1 <- clogit(Used ~ Fire_Cat + Prec_Tree + Veg_Class + L.dist2build + strata(DenID), data = data.L, method='efron')
m.2 <- clogit(Used ~ Fire_Cat + Prec_Tree  + Veg_Class+ Ave_Wet + strata(DenID), data = data.L, method='efron')
m.3 <- clogit(Used ~ Fire_Cat + Veg_Class + Ave_Wet + L.dist2build + strata(DenID), data = data.L, method='efron')
m.4 <- clogit(Used ~ Fire_Cat + Prec_Tree + Ave_Wet + L.dist2build + strata(DenID), data = data.L, method='efron')
AICc(m.0, m.1,m.2,m.3,m.4)
```
The term exp(coef) is giving the odds ratio for an increase of 1 unit in the independent variable. m.1 is now the best model. We drop Ave.Wet here instead of vegclass.

```{r fitting round 2}
m.10 <- clogit(Used ~ Fire_Cat + Prec_Tree + Veg_Class + L.dist2build + strata(DenID), data = data.L, method='efron')
m.11 <- clogit(Used ~ Fire_Cat + Veg_Class + L.dist2build + strata(DenID), data = data.L, method='efron')
m.12 <- clogit(Used ~ Fire_Cat + Prec_Tree + L.dist2build + strata(DenID), data = data.L, method='efron')
m.13 <- clogit(Used ~ Fire_Cat + Prec_Tree + Veg_Class + strata(DenID), data = data.L, method='efron')
AICc(m.10, m.11, m.12, m.13)
```
Drop Veg class (even tho only a little bit different in AIC, but more parsimonious. 
```{r likelihood ratop test}
library(lmtest)
lrtest(m.10, m.12)
```

Likelihood ratio test results a p-value of 0.07 for m.10 and m.12. Prefer m.12.

```{r fitting round 3}
m.100 <- clogit(Used ~ Fire_Cat + Prec_Tree + L.dist2build + strata(DenID), data = data.L, method='efron')
m.101 <-  clogit(Used ~ Fire_Cat + L.dist2build + strata(DenID), data = data.L, method='efron')
m.102 <- clogit(Used ~ Fire_Cat + Prec_Tree + strata(DenID), data = data.L, method='efron')
m.103 <- clogit(Used ~ Prec_Tree + strata(DenID), data = data.L, method='efron')
AICc(m.100, m.101, m.102)
```
Best model is m.100, and likeliihood ratio test proves it. 
```{r} 
lrtest(m.100, m.102)
```

Test interactions and non-linear effect among the remaining variables (only ecologically plausible ones are considered). 

```{r non-linear effect?}
# test nonlinear effect of road
data.L2 <- data.L
data.L2$L.dist2build2 <- data.L2$L.dist2build*data.L2$L.dist2build
m.10001 <- clogit(Used ~ Fire_Cat + Prec_Tree + L.dist2build +  L.dist2build2 + strata(DenID), data = data.L2, method='efron')

#test interaction terms 
m.10002 <- clogit(Used ~ Fire_Cat + Fire_Cat:Prec_Tree + L.dist2build + strata(DenID), data = data.L, method='efron')
m.10003 <- clogit(Used ~ Prec_Tree + Fire_Cat:Prec_Tree + L.dist2build + strata(DenID), data = data.L, method='efron')

AICc(m.100, m.10001, m.10002, m.10003)
```
m.100 is the best.


```{r optimal model summary}
summary(m.100)
```

## model validation

```{r hab package}
#source code https://github.com/basille/hab/blob/master/R/kfold.r 
# in the output, cor is Spearman rank correlations r_s
k_fold <- kfold(m.100, k=5, nrepet = 5, jitter = FALSE, reproducible = TRUE, details = FALSE)
rs.mean <- k_fold %>% dplyr::filter (type == "obs") %>% summarise(mean(cor))
rs.p <- t.test(k_fold %>% dplyr::filter (type == "obs") %>% dplyr::select(cor))
```

Residual deviance : `r sum(residuals(m.100, type = "deviance")^2)` on `r 913-8` degree of freedom. Over dispersion is not a problem. 

The 5-fold cross validation (sensu Boyce et al 2002) provided a mean Spearman's Rank corerltion of `r rs.mean`, P = `r rs.p$p.value`, < 0.01. 


## result plots
```{r confidence interval} 
co.table <- cbind(OR = coef(m.100), confint(m.100))
co.df <- data.frame(variable = row.names(co.table), estimate = co.table[1:nrow(co.table), 1], LL = co.table[1:nrow(co.table), 2], UL =  co.table[1:nrow(co.table), 3])
co.df$variable <- as.character(co.df$variable)

ggplot(co.df, aes(y=estimate, x=variable, ymin=LL, ymax=UL)) + geom_pointrange() + coord_flip() + ylim(-2,2)
```

The selection coefficients estimated by the conditional logistic regression are the log odds ratio for a habitat type being chosen relative to a reference habitat type (beta = 0) (there is no intercept in clogit model). As such, selection for the reference habitat occurs when the other habitat types have beta < 0. ABCD are relatively more selected than U. But only C is slightly significant, others CI overlap with 0. CI of prec_tree is high.  

Different from when random points is 10, this time it seems that fire cat C is preferred (and is significant). The results looks better with 50 random points (more fit our expectation).

``` {r predict - tree prec by fire cat}
newdata <- data.frame(DenID = factor(data.L$DenID[6]), 
                      Fire_Cat = factor(rep(c("A", "B", "C", "D", "U"), each = 100)),
                      Prec_Tree = rep(seq(from = 0, to = 100, length.out = 100), 5),
                      L.dist2build = mean(data.L$L.dist2build))

newdata <- cbind(newdata, predict(m.100, newdata = newdata, type = "lp", se = TRUE))
newdata <- within(newdata, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})

p.1 <- ggplot(newdata, aes(x = Prec_Tree, y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
    ymax = UL, fill = Fire_Cat), alpha = 0.2) + geom_line(aes(colour = Fire_Cat),
    size = 1)  + theme(legend.position = "none")


newdata <- data.frame(DenID = factor(data.L$DenID[8]), 
                      Fire_Cat = factor(rep(c("A", "B", "C", "D", "U"), each = 100)),
                      Prec_Tree = mean(data.L$Prec_Tree),
                      L.dist2build = rep(seq(from = min(data.L$L.dist2build), to = max(data.L$L.dist2build), length.out = 100), 5))

newdata <- cbind(newdata, predict(m.100, newdata = newdata, type = "lp", se = TRUE))
newdata <- within(newdata, {
  PredictedProb <- plogis(fit)
  LL <- plogis(fit - (1.96 * se.fit))
  UL <- plogis(fit + (1.96 * se.fit))
})
p.2 <- ggplot(newdata, aes(x = (10^(L.dist2build)-1), y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
    ymax = UL, fill = Fire_Cat), alpha = 0.2) + geom_line(aes(colour = Fire_Cat),
    size = 1)  + xlab("dist2build") + theme(legend.position = c(0.9, 0.2))

plot_grid(p.1, p.2)
```
